\documentclass[11pt]{article}

\usepackage[margin=0.85in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{array}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{longtable}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{microtype}

\hypersetup{
  colorlinks=true,
  linkcolor=blue!50!black,
  urlcolor=blue!50!black,
  pdftitle={Native CTS Solver Implementation Notes},
  pdfauthor={Cursor agent}
}

\lstset{
  basicstyle=\ttfamily\small,
  columns=fullflexible,
  breaklines=true,
  frame=single,
  framerule=0.2pt,
  rulecolor=\color{black!30}
}

\newcommand{\file}[1]{\nolinkurl{#1}}
\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\dd}{\mathrm{d}}

\title{Native CTS Solver Implementation Notes}
\author{AthenaK code-reading guide}
\date{May 21, 2026}

\begin{document}
\maketitle

\section{Purpose and Big Picture}

The native initial-data solver lives primarily in \file{src/z4c/id_solve.cpp} and
\file{src/z4c/id_solve.hpp}.  It solves a conformal thin sandwich (CTS) system for
four correction variables,
\[
  u = (\delta\Psi,\delta\beta^x,\delta\beta^y,\delta\beta^z),
\]
stored in \code{IDConformalThinSandwich::u\_cts}.  The correction is added to a
fixed background/free-data state stored in \code{u\_free}.  After the multigrid
solve, the code reconstructs physical ADM variables and converts them into Z4c
evolution variables.

At a high level, the flow is:
\begin{enumerate}
  \item A problem generator initializes ADM-like seed data.
  \item Z4c setup converts that seed to Z4c variables and computes initial
        diagnostics.
  \item At the start of the Z4c task list, \code{Z4c\_IDSolve} runs once
        if \code{pmy\_pack->pid\_solve} exists.
  \item The CTS solver builds free data from the current ADM variables.
  \item The multigrid driver solves for \(\delta\Psi\) and \(\delta\beta^i\).
  \item \code{ApplySolution()} writes solved physical ADM fields, converts them
        to Z4c variables, refreshes boundary/ghost zones, and recomputes ADM
        constraints.
  \item Normal hyperbolic Z4c evolution proceeds from the solved data.
\end{enumerate}

\section{File-by-File Map}

\begin{longtable}{p{0.34\linewidth} p{0.60\linewidth}}
\toprule
\textbf{File} & \textbf{Role in CTS/Z4c initial-data path} \\
\midrule
\endfirsthead
\toprule
\textbf{File} & \textbf{Role in CTS/Z4c initial-data path} \\
\midrule
\endhead
\file{src/z4c/id\_solve.hpp} &
Declares the native CTS classes and variable enums.  Defines the four solved
variables \code{ID\_CTS\_*}, all free-data slots \code{ID\_FREE\_*}, the
\code{IDCTSMultigrid} subclass, the \code{IDCTSMultigridDriver}, and the owner
class \code{IDConformalThinSandwich}. \\
\addlinespace
\file{src/z4c/id\_solve.cpp} &
Main implementation.  Contains finite-difference helpers, conformal geometry
helpers, the CTS residual operator, smoothing/defect/FAS kernels, multigrid
driver construction and solve loop, free-data construction, ADM application, and
constraint diagnostics. \\
\addlinespace
\file{src/z4c/z4c\_tasks.cpp} &
Queues \code{IDConformalThinSandwich::SolveTask} as \code{Z4c\_IDSolve} in the
\code{Task\_Start} phase after \code{Z4c\_Recv}.  Also shows the evolution task
sequence that follows: copy, RHS, boundary RHS, RK update, restriction, send/recv,
physical BCs, prolongation, algebraic constraints, ADM conversion, diagnostics,
tracking, CCE, and horizon finding. \\
\addlinespace
\file{src/z4c/z4c.cpp} &
Constructs Z4c options and boundary buffers.  Maps \code{z4c/spatial\_order} to
\code{opt.fd\_stencil}: 2nd order \(\to 2\), 4th order \(\to 3\), 6th order
\(\to 4\).  Allocates \code{pbval\_u} and \code{pbval\_weyl}. \\
\addlinespace
\file{src/z4c/z4c\_adm.cpp} &
Implements \code{ADMToZ4c}, \code{Z4cToADM}, and \code{ADMConstraints}.  This is
the conversion bridge between solved ADM data and Z4c evolution arrays. \\
\addlinespace
\file{src/pgen/z4c\_critical\_kerr\_formation.cpp} &
Problem generator for the compact-support conformal-metric seed.  Builds
\(\bar{\gamma}_{ij}\), a trace-free time-derivative/extrinsic-curvature seed,
zero shift, and chosen lapse; then converts ADM to Z4c before the CTS task runs. \\
\addlinespace
\file{src/multigrid/multigrid.hpp} &
Base multigrid data structures.  Defines \code{Multigrid}, \code{MGOctet}, task
IDs, storage for \code{u}, \code{src}, \code{def}, \code{coeff}, and generic
restriction/prolongation/smoothing interfaces. \\
\addlinespace
\file{src/multigrid/multigrid.cpp} &
Generic multigrid operations: load finest data/source/coefficient arrays,
restrict coefficients, retrieve solved data, restrict residuals, prolong and
correct, compute defect norms, root-grid transfers. \\
\addlinespace
\file{src/multigrid/multigrid\_driver.cpp} &
Base multigrid driver: hierarchy setup, V-cycle and FMG orchestration, root-grid
boundary treatment, octet/coarse-fine bridge support, default multigrid boundary
condition selection. \\
\addlinespace
\file{src/multigrid/multigrid\_tasks.cpp} &
Task-list wiring for multigrid cycles: boundary receive/send, physical boundary
fill, fine-coarse ghost fill, red/black smoothing, restriction, prolongation, and
cleanup dependencies. \\
\addlinespace
\file{src/utils/finite\_diff.hpp} &
Generated finite-difference stencils used by CTS and Z4c.  The template
parameter \code{NGHOST=2,3,4} corresponds to 2nd, 4th, and 6th-order centered
operators. \\
\addlinespace
\file{inputs/z4c/gw\_collapse/z4c\_critical\_kerr\_formation.athinput} &
Representative input file.  Sets mesh boundaries, \code{z4c/spatial\_order},
\code{id\_solve} controls, and the critical-Kerr seed parameters. \\
\bottomrule
\end{longtable}

\section{Objects and Data Arrays}

\subsection{Solved Variables}

\file{src/z4c/id\_solve.hpp} defines:
\begin{lstlisting}
enum IDCTSUVar {
  ID_CTS_PSI,
  ID_CTS_BETAX,
  ID_CTS_BETAY,
  ID_CTS_BETAZ,
  ID_CTS_NVAR
};
\end{lstlisting}

The multigrid unknown \code{u\_cts} stores corrections to a base conformal factor
and base shift.  The actual fields used by the residual are assembled with
\code{TotalU()}:
\[
  \Psi = \code{ID\_FREE\_BASE\_PSI} + u_{\Psi}, \qquad
  \beta^i = \code{ID\_FREE\_BASE\_BETA}^i + u_{\beta^i}.
\]
In the current critical-Kerr seed, the base values are \(\Psi=1\) and
\(\beta^i=0\), so the solver starts from no correction.

\subsection{Free-Data Variables}

\code{IDCTSFreeVar} includes:
\begin{itemize}
  \item \code{ID\_FREE\_GXX ... ID\_FREE\_GZZ}: conformal metric
        \(\bar{\gamma}_{ij}\).
  \item \code{ID\_FREE\_GDOTXX ... ID\_FREE\_GDOTZZ}: contravariant trace-free
        tensor \(\bar{u}^{ij}_{TF}\).  The names say \code{GDOT}, but the comment
        in the header clarifies the actual stored quantity.
  \item \code{ID\_FREE\_K}: trace \(K\).
  \item \code{ID\_FREE\_DKX ... ID\_FREE\_DKZ}: \(\bar{D}^i K\).
  \item \code{ID\_FREE\_ALPHA}: physical/free ADM lapse copied from
        \code{admvars.alpha} and used by the CTS operator.
  \item \code{ID\_FREE\_E}: Eulerian energy density source.
  \item \code{ID\_FREE\_PX/PY/PZ}: conformally raised momentum source
        \(\bar{\gamma}^{ij}S_j\).
  \item \code{ID\_FREE\_SOURCE}: extra scalar source in the Hamiltonian equation.
  \item \code{ID\_FREE\_MASK}: active-cell mask for the residual and smoother.
  \item \code{ID\_FREE\_BASE\_PSI}, \code{ID\_FREE\_BASE\_BETAX/BETAY/BETAZ}:
        base values to which corrections are added.
\end{itemize}

\section{Problem-Generator Seed Path}

\file{src/pgen/z4c\_critical\_kerr\_formation.cpp} is the current compact-support
test seed.  The important functions are:

\begin{itemize}
  \item \code{ReadOptions()}: reads \code{problem/*} options such as amplitude,
        support radius, helicity, phase, lapse, and radial coefficients.
  \item \code{RadialProfileAndDerivative()}: constructs a compact bump profile
        and radial derivative using Legendre modes.
  \item \code{FillCriticalKerrFreeDataADM()}: fills ADM fields on active and
        ghost cells.  It constructs a regular \(\ell=m=2\) solid-harmonic tensor
        perturbation, builds \(q_{ij}=\delta_{ij}+h_{ij}+h^2_{ij}/2\), rescales it
        to unit determinant, and stores it as an ADM spatial metric seed.  It also
        builds a trace-free \(\dot{u}_{ij}\)-like tensor and stores
        \(K_{ij}=0.5\,\dot{u}_{ij}\), with zero shift and configured lapse.
  \item \code{ConvertADMToZ4cAndConstraints()}: dispatches to
        \code{ADMToZ4c<2/3/4>} based on \code{pz4c->opt.fd\_stencil}, computes
        ADM constraints, then calls \code{Z4cToADM()} so the ADM view is
        consistent with the Z4c storage.
  \item \code{ProblemGenerator::UserProblem()}: checks that \code{z4c}, \code{adm},
        and \code{id\_solve} exist, fills the seed, optionally sets a
        pre-collapsed lapse, converts to Z4c, and returns.
\end{itemize}

\section{Task Scheduling and Handoff Timing}

The solve is inserted in \file{src/z4c/z4c\_tasks.cpp}.  In
\code{Z4c::QueueZ4cTasks()}, the start phase is:
\begin{lstlisting}
Z4c_Recv     = Z4c::InitRecv
Z4c_IRecvW   = Z4c::InitRecvWeyl
Z4c_IDSolve  = IDConformalThinSandwich::SolveTask, depends on Z4c_Recv
\end{lstlisting}

This means Z4c boundary receives are posted before the ID solve.  The solve is
still before the hyperbolic RHS and RK update.  After the solve, regular evolution
continues through \code{CopyU}, \code{CalcRHS}, \code{Z4cBoundaryRHS},
\code{ExpRKUpdate}, mesh communication, physical BCs, prolongation, algebraic
constraint enforcement, and \code{Z4cToADM}.

Inside \code{IDConformalThinSandwich::SolveTask()}:
\begin{enumerate}
  \item Optionally write ``before'' constraint diagnostics.
  \item Call \code{IDCTSMultigridDriver::Solve()}.
  \item If a solution was accepted, call \code{RefreshZ4cBoundariesAfterSolve()}.
  \item Recompute constraints on the refreshed solution.
  \item Optionally write ``after'' diagnostics.
\end{enumerate}

The refresh step is important.  \code{ApplySolution()} writes Z4c data, but ghost
zones and physical boundaries need to be made consistent before evolution and
before final diagnostics.

\section{Free-Data Construction}

\code{IDConformalThinSandwich::BuildFreeData()} converts the current ADM fields
into the free-data arrays consumed by the CTS residual.

\subsection{Metric, Lapse, Shift, and Trace}

For every active and ghost cell:
\begin{itemize}
  \item \(\bar{\gamma}_{ij}\) is copied from \code{admvars.g\_dd}.
  \item The inverse metric is computed locally.
  \item \(K=\gamma^{ij}K_{ij}\) is computed from \code{admvars.vK\_dd}.
  \item \code{ID\_FREE\_ALPHA} is copied from \code{admvars.alpha}.  The current
        implementation treats this as the physical/free ADM lapse throughout the
        solve; it is not rescaled as a separate conformal lapse at handoff.
  \item The base conformal factor is set to 1.
  \item The base shift is copied from \code{admvars.beta\_u}.
  \item The correction variables in \code{u\_cts} are reset to zero.
\end{itemize}

\subsection{Matter Sources and Masks}

If \code{pmy\_pack->ptmunu} exists, the code reads Eulerian energy density
\code{tmunu.E} and raises the covariant momentum \code{S\_d} using the conformal
metric inverse.  Otherwise \(E=0\) and \(p^i=0\).  With this storage convention,
the momentum residual term \(\Psi^6 P^i\) corresponds to the usual
\(\Psi^{10}j^i\) physical-momentum source after the conformal raising is
accounted for.

The mask controls where the residual and smoother are active.  If
\code{id\_solve/mask\_horizon\_defect=true}, cells inside
\code{horizon\_mask\_radius} get \code{ID\_FREE\_MASK=0}.  If
\code{id\_solve/fill\_horizon\_junk=true}, source terms inside
\code{horizon\_fill\_radius} are adjusted by \code{FillHorizonJunk()}.

\subsection{\(\bar{u}^{ij}_{TF}\) and \(\bar{D}^i K\)}

\code{BuildGammaDotAndDK<NGHOST>()} computes:
\begin{itemize}
  \item \(\bar{D}^i K = \bar{\gamma}^{ij}\partial_j K\), stored in
        \code{ID\_FREE\_DKX/DKY/DKZ}.
  \item The trace-free extrinsic-curvature seed \( \hat{A}^{ij} \) by raising the
        trace-free part of \(K_{ij}\) using the conformal metric.
  \item The conformal longitudinal shift operator \((\bar{L}\beta)^{ij}\).
  \item The stored tensor
        \[
          \bar{u}^{ij}_{TF} = 2\alpha\hat{A}^{ij} - (\bar{L}\beta)^{ij}.
        \]
\end{itemize}

The halo width needed here follows the compact \(D_j\hat{A}^{ij}\) path: the
stored trace-free tensor must be available far enough into the ghost zone for
the later centered derivative in the momentum residual.

\section{CTS Residual Operator}

\code{CTSOperator<NGHOST>()} evaluates the four residual equations at one cell.
Its inputs are:
\begin{itemize}
  \item \code{u}: current multigrid correction variables.
  \item \code{free}: coefficient/free-data array.
  \item \code{idx}: inverse grid spacing.
  \item \code{fd\_stencil}: base derivative stencil selector.
\end{itemize}

\subsection{Geometry Helpers}

\begin{itemize}
  \item \code{MetricInverse()} computes \(\bar{\gamma}^{ij}\).
  \item \code{Christoffel<NGHOST>()} computes \(\bar{\Gamma}^i{}_{jk}\) using
        centered finite differences of the conformal metric.
  \item \code{RicciScalar<NGHOST>()} computes \(\bar{R}\) from first and second
        finite differences of \(\bar{\gamma}_{ij}\).
  \item \code{DxTotal()}, \code{DxxTotal()}, and \code{DxyTotal()} differentiate
        base-plus-correction quantities.
\end{itemize}

\subsection{\(\hat{A}^{ij}\)}

\code{AHatUU<NGHOST>()} computes:
\[
  \hat{A}^{ij} =
  \frac{\bar{u}^{ij}_{TF} + (\bar{L}\beta)^{ij}}{2\alpha}.
\]
\code{DxAHatUU()} applies the chain rule to differentiate \(\hat{A}^{ij}\)
directly at the cell center.  The derivative is expanded into first and second
finite differences of the primitive fields \(\beta^i\), \(\bar{\gamma}_{ij}\),
\(\bar{u}^{ij}_{TF}\), and \(\alpha\).  These derivatives use the existing
\file{src/utils/finite\_diff.hpp} operators: \code{Dx}, \code{Dxx}, \code{Dxy},
and the base-plus-correction wrappers \code{DxTotal}, \code{DxxTotal}, and
\code{DxyTotal}.  This compact form is the only implemented
\(D_j\hat{A}^{ij}\) path.

\subsection{Hamiltonian Residual}

The implemented scalar residual is:
\[
  \bar{D}^2\Psi
  - \frac{1}{8}\bar{R}\Psi
  + \frac{1}{8}\hat{A}_{ij}\hat{A}^{ij}\Psi^{-7}
  + 2\pi E\Psi^5
  - \frac{1}{12}K^2\Psi^5
  - S = 0.
\]
Here \(S\) is \code{ID\_FREE\_SOURCE}.  In vacuum critical-Kerr tests,
\(E=S=0\).

\subsection{Momentum Residual}

For each component \(i\), the implemented residual is:
\[
  2\bar{D}_j\hat{A}^{ij}
  - \frac{4}{3}\Psi^6\bar{D}^iK
  - 16\pi\Psi^6 P^i = 0.
\]
\(\bar{D}_j\hat{A}^{ij}\) is assembled from \(\partial_j\hat{A}^{ij}\) plus
Christoffel terms.

\subsection{Stencil and Halo Selection}

\code{z4c/spatial\_order} maps to \code{fd\_stencil} as:
\[
  2 \mapsto 2,\qquad 4 \mapsto 3,\qquad 6 \mapsto 4.
\]
\code{NGHOST=2,3,4} in \file{finite\_diff.hpp} gives 2nd, 4th, and 6th-order
centered derivative formulas.

The \(D_j\hat{A}^{ij}\) path is compact: it differentiates the expanded
expression directly and uses the base finite-difference stencil.  Because the
current free-data setup computes \(\bar{u}^{ij}_{TF}\) from ADM data before the
solve, that prepass fills a halo of width \(\code{fd\_stencil}-1\) for
\(\bar{u}^{ij}_{TF}\).  Computing that halo also uses centered derivatives, so
the free-data prepass requires enough mesh ghost cells for the prepass stencil
plus the stored \(\bar{u}^{ij}_{TF}\) halo.  For the current 4th-order
critical-Kerr tests this is \code{mesh/nghost=4}.  In general the constructor
requires
\[
  \code{mesh/nghost} \ge 2(\code{fd\_stencil}-1),
\]
which is a CTS free-data prepass requirement.  The multigrid correction-variable
halo separately requires
\[
  \code{id\_solve/mg\_nghost} \ge \code{fd\_stencil}.
\]

The fine CTS operator order is used only on the finest MeshBlock MG level.
Coarser MeshBlock levels and all root-grid levels use
\code{id\_solve/mg\_coarse\_fd\_stencil}; the default is \code{2} when the fine
operator is higher order.  This keeps high-order accuracy on the final residual
while making coarse-level FAS operators less halo hungry and less sensitive to
coarse coefficient interpolation.  The octet bridge still uses the independent
\code{id\_solve/octet\_fd\_stencil} control and remains guarded for sixth-order
SMR.

\section{Smoothing Kernel}

The smoother is in \code{SmoothImpl()} and is called by
\code{IDCTSMultigrid::SmoothPack()}.  The driver reads
\code{id\_solve/smoother}; the default is the original red-black, pointwise,
diagonally preconditioned nonlinear relaxation:
\[
  u_v \leftarrow u_v - \omega\,
  \frac{\mathcal{F}_v(u)-{\rm src}_v}{D_v}.
\]
Here \(D_v\) is the approximate diagonal coefficient returned by
\code{CTSOperator()}.  The scalar equation uses a diagonal built from the
centered Laplacian coefficient and local nonlinear derivatives.  The shift
equations use a diagonal approximation to the vector Laplacian-like principal
part divided by \(\alpha\).

This diagonal mode has the following properties:
\begin{itemize}
  \item The update writes in-place to the live correction view and is red-black
        colored.
  \item It does not solve a coupled local \(4\times4\) Newton system for
        \((\Psi,\beta^i)\).
  \item It ignores off-diagonal Jacobian couplings between \(\Psi\) and
        \(\beta^i\), and between shift components.
  \item It is therefore closer to an under-relaxed diagonal Newton/Jacobi-GS
        smoother.
\end{itemize}

The optional \code{smoother = newton\_gs} path applies a coupled local Newton
update at each active red/black point.  At a fixed cell, it forms the local
residual
\[
  R_v = \mathcal{F}_v(u)-{\rm src}_v
\]
for \(v=(\Psi,\beta^x,\beta^y,\beta^z)\), builds a scale-aware \(4\times4\)
Jacobian by finite differencing the full local CTS operator with respect to the
four cell-centered unknowns, solves
\[
  J_{vw}\,\delta u_w = -R_v,
\]
and applies a residual-checked damped update.  The finite-difference Jacobian
uses a read-only local perturbed view of the solution, so neighboring threads
never observe a temporary perturbation of the shared grid.  The local linear
solve scales rows and columns before pivoting; if the solve is singular or
ill-conditioned, the code falls back to the diagonal update at that cell.

The Newton-Gauss-Seidel option is closer to the relaxation described by East
et al. (arXiv:1208.3473), because it includes the local coupling among \(\Psi\)
and the three shift components.  It is still damped by \(\omega\).  The
residual-checked line search makes it much harder for this mode to take
destructive coupled Newton directions, but it remains more expensive than the
diagonal smoother because it finite-differences and solves a local \(4\times4\)
system at each active point.

The smoothing update mode is now \code{smoother\_update=frozen\_view}.  Before
each smoothing visit, the active MG level is copied into a dedicated scratch
view.  \code{CTSOperator()}, the diagonal update, the \code{newton\_gs} Jacobian,
and the line-search residual all read from this frozen view, while accepted
updates write to the live \code{u} view.  This removes same-color read/write
races from mixed derivatives and high-order centered stencils without replacing
the current two-color task structure with a much larger multi-color schedule.

Properties common to both smoother modes:
\begin{itemize}
  \item The update is active only where \code{ID\_FREE\_MASK >= 0.5}.
  \item MeshBlock smoothing and root-octet smoothing use the same smoother mode.
  \item MeshBlock/root smoothing uses a dedicated \code{DualArray5D} frozen
        scratch view; octet smoothing uses a local frozen copy of the octet data.
  \item Non-finite proposed updates are rejected.
  \item \code{ngs\_max\_update} gives an optional absolute update cap, and
        \code{smoother\_max\_update\_fraction} caps updates relative to the local
        total-field scale.
  \item After each accepted or fallback point update it floors \(\Psi\) at
        \(10^{-8}\).
\end{itemize}

\subsection{Smoother Production Safeguards}

The coupled Newton-Gauss-Seidel smoother is the more faithful algorithmic target
for a nonlinear FAS CTS solve, but it should still earn default status through
wall-clock validation rather than through formulation alone.  The implemented
production safeguards are:

\begin{enumerate}
  \item For \code{newton\_gs}, after solving \(J\delta u=-R\), the code tries
        \(\lambda\omega\delta u\) with \(\lambda=1,1/2,1/4,\ldots\).  It accepts
        the first candidate that reduces the local residual norm and preserves
        \(\Psi>0\).
  \item If no Newton candidate is acceptable, or if the scaled \(4\times4\) solve
        is singular, the point falls back to the diagonal update.
  \item The local solve scales columns by the local total-field magnitude and
        scales rows before pivoting, so \(\Psi\) and \(\beta^i\) are not treated
        as if they had identical natural magnitudes.
  \item Both smoothers use finite-value guards and optional update limiting.
  \item Per-V-cycle diagnostics count accepted updates, limited updates,
        backtracked updates, diagonal fallbacks, singular local solves,
        non-finite proposals, rejected updates, \(\Psi\) floors, and maximum
        update.  These counters are reduced across MPI ranks and printed when
        \code{show\_smoother\_stats=true}.
\end{enumerate}

If the coupled smoother becomes consistently faster in defect reduction per
wall-clock time, then it becomes a good candidate for the default.  Until then,
the diagonal smoother is less complete mathematically but remains the more
predictable production default.

The default/used solver controls are read in
\code{IDCTSMultigridDriver::IDCTSMultigridDriver()}:
\begin{itemize}
  \item \code{id\_solve/omega}: relaxation factor.
  \item \code{id\_solve/smoother}: \code{diagonal} (default) or
        \code{newton\_gs}.
  \item \code{id\_solve/smoother\_update}: currently \code{frozen\_view}.
  \item \code{id\_solve/ngs\_iterations}: number of local Newton updates per
        smoothing visit when \code{smoother = newton\_gs}.
  \item \code{id\_solve/ngs\_jacobian\_eps}: relative perturbation used to
        finite-difference the local \(4\times4\) Jacobian, scaled by the local
        total-field magnitude.
  \item \code{id\_solve/ngs\_max\_update}: optional absolute limiter on each
        component update; \code{0} disables this limiter.
  \item \code{id\_solve/ngs\_line\_search\_steps}: maximum number of local
        Newton backtracking attempts.
  \item \code{id\_solve/ngs\_line\_search\_min}: minimum accepted line-search
        damping factor.
  \item \code{id\_solve/smoother\_max\_update\_fraction}: optional update cap
        as a fraction of the local total-field scale.
  \item \code{id\_solve/show\_smoother\_stats}: print MPI-reduced smoother
        counters after each V-cycle.
  \item \code{id\_solve/npresmooth}, \code{id\_solve/npostsmooth}: V-cycle
        smoothing counts.
  \item \code{id\_solve/niteration}: fixed V-cycle budget when
        \code{threshold < 0}.
  \item \code{id\_solve/threshold}: convergence threshold when nonnegative.
  \item \code{id\_solve/full\_multigrid}: optional FMG solve.
  \item \code{id\_solve/keep\_best\_solution} and
        \code{id\_solve/stop\_on\_defect\_increase}: guards for unstable or
        over-relaxed runs.
\end{itemize}

\section{Multigrid Solve Lifecycle}

\code{IDCTSMultigridDriver::Solve()} does the following:
\begin{enumerate}
  \item \code{PrepareForAMR()} constructs root-grid and octet bookkeeping.
  \item Guards reject adaptive AMR and currently reject sixth-order CTS on SMR.
  \item \code{owner\_->BuildFreeData()} builds \code{u\_free} and resets
        \code{u\_cts}.
  \item \code{mglevels\_->LoadFinestData(owner\_->u\_cts, ...)} loads the
        correction initial guess.
  \item \code{LoadSource(owner\_->u\_defect, ..., 0.0)} sets the source to zero.
        This makes the problem \(F(u)=0\) on the finest level; FAS modifies source
        arrays on coarser levels.
  \item \code{LoadCoefficients(owner\_->u\_free, ...)} loads free-data
        coefficients into the multigrid coefficient arrays and fills their
        same-level and physical halos on the finest MeshBlock MG level.
  \item \code{RestrictCoefficients()} restricts free-data arrays down the block
        hierarchy and fills coefficient halos after each coarse restriction.
        Root coefficients are filled with periodic copies or zero-gradient
        physical copies, not with correction-variable \code{mg\_zerofixed}
        sign flips.
  \item \code{SetupMultigrid()} initializes levels, tasks, root grids, and octets.
  \item \code{TransferCoefficientsFromBlocksToRoot()} populates root and octet
        coefficient storage.
  \item Initial defect norms are computed with \code{CalculateDefectNorm()}.
  \item The code runs FMG, thresholded V-cycles, or fixed-count V-cycles.
  \item If accepted, \code{RetrieveResult()} copies the solved correction back
        into \code{owner\_->u\_cts}.
  \item \code{owner\_->ApplySolution()} reconstructs ADM and Z4c data.
\end{enumerate}

\section{Restriction, Prolongation, and Defect Norms}

\file{src/multigrid/multigrid.cpp} provides the generic transfer operations:
\begin{itemize}
  \item \code{RestrictPack()}: computes the defect and restricts it to the
        coarser source.  It also restricts the current solution.
  \item \code{ProlongateAndCorrectPack()}: computes a coarse-grid correction and
        prolongates it to the next-finer solution.
  \item \code{FMGProlongatePack()}: direct FMG prolongation.
  \item \code{CalculateDefectNorm()}: computes L1/L2/Linf norms.  The current L1
        and L2 paths multiply by cell volume, so the L2 norm is volume-normalized
        after the driver scale factor is applied.
\end{itemize}

The V-cycle task lists in \file{src/multigrid/multigrid\_tasks.cpp} perform
boundary exchange before smoothing/restriction/prolongation, then call
\code{PhysicalBoundary()} and \code{FillFCBoundary()} before using ghost cells.

\section{Boundary Conditions}

\subsection{CTS Multigrid Boundary Conditions}

The base multigrid constructor in \file{src/multigrid/multigrid\_driver.cpp}
initializes multigrid boundary conditions from the mesh:
\begin{itemize}
  \item Periodic mesh faces remain periodic.
  \item Every non-periodic mesh face defaults to \code{mg\_zerofixed} in the
        generic base driver.
\end{itemize}

The CTS driver then overrides non-periodic faces according to
\code{id\_solve/mg\_bc}.  The current default is \code{zerograd}, so the critical-
Kerr input file uses \code{mg\_zerograd} on all six physical faces because all
mesh faces are \code{outflow}.  Set \code{mg\_bc=zerofixed} to recover the older
Dirichlet-zero correction boundary.

The physical boundary fill in \code{MultigridDriver::PhysicalBoundary()} and
\code{MGRootBoundary()} implements:
\begin{itemize}
  \item \code{mg\_zerofixed}: odd reflection of the correction variable,
        approximately enforcing zero Dirichlet value at the boundary.
  \item \code{mg\_zerograd}: even reflection, enforcing zero normal derivative.
  \item \code{periodic}: periodic copy.
  \item \code{mg\_multipole}: multipole potential boundary support, inherited
        from gravity multigrid, but not currently exposed by \code{id\_solve}
        input parameters.
\end{itemize}

Unlike the gravity driver, CTS currently exposes only \code{zerograd} and
\code{zerofixed} through \code{id\_solve/mg\_bc}; multipole MG boundaries are not
used by the CTS solve.

\subsection{Coefficient and Coarse/Fine Boundaries}

The free-data coefficients are loaded into multigrid coefficient arrays and
restricted down levels.  For SMR/octet handling:
\begin{itemize}
  \item \code{TransferCoefficientsFromBlocksToRoot()} copies block coefficients
        to root cells or refined octets.
  \item Refined coefficient octets are restricted upward/downward as needed.
  \item Same-level octet neighbors copy boundary data.
  \item Coarse/fine octet boundaries use interpolation/prolongation helpers.
  \item Physical coefficient boundaries are filled by
        \code{ApplyCoefficientPhysicalBoundaries()} using Lagrange extrapolation
        for non-mask coefficients and direct copy for masks.
\end{itemize}

\subsection{Z4c Boundary Conditions After Handoff}

After \code{ApplySolution()} has written Z4c variables, the solve task refreshes
Z4c boundaries:
\begin{enumerate}
  \item \code{RestrictU()}
  \item \code{SendU()}
  \item \code{ClearSend()}
  \item \code{ClearRecv()}
  \item \code{RecvU()}
  \item \code{Z4cBoundaryRHS()}
  \item \code{ApplyPhysicalBCs()}
  \item \code{Prolongate()}
  \item \code{InitRecv()} again if evolution will continue
\end{enumerate}

\code{Z4c::ApplyPhysicalBCs()} calls \code{pbval\_u->Z4cBCs(...)} unless the mesh
is strictly periodic, then optional user boundary conditions from the pgen.
Thus, CTS solve boundary conditions and Z4c evolution boundary conditions are
separate:
\begin{itemize}
  \item CTS solves corrections with multigrid correction BCs.
  \item The resulting Z4c fields are then put through normal Z4c ghost exchange
        and physical BC machinery before evolution.
\end{itemize}

\section{Applying the Solution}

\code{IDConformalThinSandwich::ApplySolution()} is the handoff from elliptic
variables to evolution variables.

At each active cell:
\begin{itemize}
  \item Compute \(\Psi = \code{base\_psi}+\delta\Psi\), then \(\Psi^4\).
  \item Set physical ADM metric
        \[
          \gamma_{ij} = \Psi^4 \bar{\gamma}_{ij}.
        \]
  \item Set \code{admvars.psi4} from \(\Psi^4\det(\bar{\gamma})^{1/3}\).
  \item Copy lapse from \code{ID\_FREE\_ALPHA}.
  \item Set physical shift
        \[
          \beta^i = \beta^i_{\rm base}+\delta\beta^i.
        \]
  \item Recompute \(\hat{A}^{ij}\) from the solved shift.
  \item Lower \(\hat{A}^{ij}\) with the conformal metric and reconstruct
        \[
          K_{ij} = \Psi^{-2}\hat{A}_{ij}
                 + \frac{1}{3}\gamma_{ij}K.
        \]
\end{itemize}

Then \code{ApplySolution()} calls:
\begin{enumerate}
  \item \code{ADMToZ4c<fd>()}: populate Z4c evolution variables.
  \item \code{Z4cToADM()}: make the ADM diagnostic view consistent with Z4c.
  \item \code{ADMConstraints<fd>()}: compute constraint diagnostics on the active
        region.
\end{enumerate}

Finally, \code{SolveTask()} performs the boundary refresh described above and
recomputes constraints one more time on the boundary-consistent state.

\section{Current Critical-Kerr Input Choices}

In \file{inputs/z4c/gw\_collapse/z4c\_critical\_kerr\_formation.athinput}, the
current representative setup is:
\begin{itemize}
  \item Mesh BCs: \code{outflow} on all faces.
  \item ADM dynamic: \code{true}.
  \item Z4c spatial order: \code{4}, so \code{fd\_stencil=3}.
  \item CTS enabled and solve-once.
  \item \code{omega=0.02} in the file, though convergence sweeps found
        larger values useful for short second-order diagonal-smoother runs.
  \item \code{smoother=diagonal} by default.  Set
        \code{smoother=newton\_gs} to use the coupled local
        Newton-Gauss-Seidel relaxation.
  \item \code{smoother\_update=frozen\_view}: read operator values from a
        scratch copy during each smoothing visit and write updates to the live
        correction view.
  \item \code{ngs\_line\_search\_steps=8},
        \code{ngs\_line\_search\_min=1.0e-4}, and
        \code{smoother\_max\_update\_fraction=0.25}: local update safeguards.
  \item \code{threshold=-1.0}: fixed V-cycle count mode.
  \item \code{niteration=1} in the input file; convergence tests override this.
  \item \code{npresmooth=2}, \code{npostsmooth=2}.
  \item \code{show\_smoother\_stats=true}: print per-V-cycle smoother counters.
  \item \code{mg\_nghost=4}: order-matched for current 4th-order critical-Kerr
        CTS tests.  For 6th order with the current free-data prepass, use
        \code{mesh/nghost=6} and \code{id\_solve/mg\_nghost=6}.
  \item \code{mg\_coarse\_fd\_stencil=2}: use the second-order CTS operator on
        coarse MeshBlock/root levels while keeping the fine-grid operator at the
        configured Z4c order.
  \item \code{mg\_bc=zerograd}: use zero-normal-gradient MG boundaries for CTS
        correction variables on non-periodic faces.
  \item \code{reject\_worse=false}, \code{keep\_best\_solution=true}.
\end{itemize}

\section{Current Smoother Validation Snapshot}

The production-smoother validation runs are recorded under
\file{analysis/critical\_kerr\_convergence/smoother\_comparison}.  All runs used
MPI with four ranks or fewer.  The current snapshot is:
\begin{itemize}
  \item At \(32^3\), second order, 10 V-cycles, the diagonal smoother was stable
        for \(\omega=0.02,0.05,0.08\).  The new residual-checked
        \code{newton\_gs} path was stable for \(\omega=0.005,0.02,0.05\).
  \item At \(64^3\), second order, 40 V-cycles, \code{diagonal},
        \(\omega=0.05\), reduced the total CTS defect by a factor
        \(7.10\times10^{-2}\) in about \(152\) seconds.
  \item The matching \code{newton\_gs}, \(\omega=0.05\), run reduced the total
        CTS defect by a factor \(1.46\times10^{-2}\), but took about \(525\)
        seconds.  It therefore reached a lower final defect but was slower in
        defect reduction per wall-clock time for this case.
  \item At \(32^3\), fourth order with \code{nghost=4}, both smoothers completed
        the short 5-cycle smoke test without non-finite updates.
  \item A short \code{mg\_bc=zerofixed} comparison also remained bounded, though
        \code{mg\_bc=zerograd} remains the less intrusive default.
\end{itemize}

The smoother counters showed no singular local solves, non-finite proposals,
fallbacks, or \(\Psi\) floors in these validation cases.  That means the new
safeguards are not masking a frequent failure in this regime, but they remain
important for stronger free data, coarser grids, and FAS coarse-level states.

\section{Known Implementation Caveats}

\begin{itemize}
  \item The default smoother remains the diagonal approximate Newton smoother
        for continuity with previous runs.  The optional \code{newton\_gs}
        smoother now has local residual-checked globalization and scale-aware
        linear algebra, but it is still more expensive because it builds a
        finite-difference Jacobian at every active point.
  \item Frozen-view smoothing removes same-color races for the current stencil
        family.  It may be a weaker smoother than a fully stencil-aware
        multi-color Gauss-Seidel schedule, so multi-color remains a possible
        future optimization if frozen-view convergence is too slow.
  \item The matter-source convention in the residual uses \(E\Psi^5\) and
        \(P^i\Psi^6\), where \(P^i=\bar{\gamma}^{ij}S_j\) is the stored
        conformally raised momentum source.  In vacuum this is irrelevant, but
        matter runs should preserve this convention consistently.
  \item \code{ID\_FREE\_ALPHA} is currently the physical/free ADM lapse used
        consistently by \code{BuildFreeData()}, \code{CTSOperator()}, and
        \code{ApplySolution()}.  A future conformal-lapse convention would need
        explicit conversion both when building free data and when applying the
        solution.
  \item The CTS driver exposes \code{id\_solve/mg\_bc=zerograd|zerofixed}.  This
        controls only the CTS multigrid correction-variable boundary; after
        handoff, Z4c fields still use the normal Z4c/mesh boundary machinery.
  \item The compact \(D_j\hat{A}^{ij}\) path should be validated by convergence
        tests whenever changing order, boundary conditions, or free data.
  \item Coarse MeshBlock/root levels intentionally default to lower-order CTS
        operators.  This should not limit fine-grid convergence if the fine
        residual is solved tightly, but it should be included in validation
        reports.
  \item Sixth-order CTS on SMR is explicitly guarded off because the current octet
        coarse/fine bridge is not widened enough for the required stencil.
\end{itemize}

\section{Where to Start Reading}

A good code-reading order is:
\begin{enumerate}
  \item \file{src/z4c/id\_solve.hpp}: understand variable enums and class layout.
  \item \file{src/pgen/z4c\_critical\_kerr\_formation.cpp}: see how the seed data
        are constructed.
  \item \file{src/z4c/z4c\_tasks.cpp}: see where the solve is queued.
  \item \file{src/z4c/id\_solve.cpp}, starting with:
        \code{IDConformalThinSandwich::SolveTask()},
        \code{IDCTSMultigridDriver::Solve()},
        \code{BuildFreeData()},
        \code{CTSOperator()},
        \code{SmoothImpl()}, and \code{ApplySolution()}.
  \item \file{src/multigrid/multigrid\_driver.cpp} and
        \file{src/multigrid/multigrid\_tasks.cpp}: see how V-cycles, red/black
        smoothing, and boundary exchanges are scheduled.
  \item \file{src/z4c/z4c\_adm.cpp}: see the ADM/Z4c conversion and constraint
        diagnostics.
\end{enumerate}

\end{document}
